---
title: "GO analysis"
output:
  pdf_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---


#Load libraries, DB and the main table

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

#BiocManager::install("clusterProfiler")
library(clusterProfiler)

#BiocManager::install("org.Hs.eg.db")
library(org.Hs.eg.db)

library(ggplot2)
library(tidyr)
library(reshape2)


proteinGroups_table <- read.table("/Users/ananth/Documents/MaxQuant_Bechmarking/Human/Bin_intensity_analysis/Gene_distribution_in_organs-GeneNames-Median_bin_values.txt",
                                  header = TRUE,
                                  stringsAsFactors = FALSE)
#proteinGroups_table <- proteinGroups_table[,-c(35,36,37,38)]
```


The proteinGroups_table is a ppb normalized intensity matrix by sample.

The following script transform the values in log10 and perform a median normalization across samples


```{r}
#proteinGroups_table_tmp <- proteinGroups_table[,c(-1,-2, -3)]
proteinGroups_table_tmp <- proteinGroups_table[,c(-1,-2)]
median_values <- apply(proteinGroups_table_tmp, 1, function(x) median(na.omit(x)))
proteinGroups_norm <- proteinGroups_table_tmp/median_values
proteinGroups_norm[is.na(proteinGroups_norm)] <- 0
proteinGroups_norm <- cbind(proteinGroups_table[,c(1,2)], proteinGroups_norm)
####rm(median_values, proteinGroups_table_tmp)
proteinGroups_mean <- apply(proteinGroups_norm[, -c(1,2)], 1, mean)

```

Each gene-tissue is classified in one of 3 categories: 
   - "Tissue enriched" at least 10-fold higher intesnity compared to average value and must be unique high value.
   - "Group enriched" at least 10-fold higher intesnity compared to average value in a group of 2-7 tissues compared to al other tissues
   - "Mixed" when not fulfil any of the above categories (they are discarded)

```{r}
Elevated_genes <- data.frame(
  "Gene ID" = NA,
  "Tissue" = NA,
  "Gene name" = NA,
  "Category" = NA
)

for (i in seq(dim(proteinGroups_norm)[1])){
  get_sign_values <- which(proteinGroups_norm[ i, c(-1,-2)] > proteinGroups_mean[i]*2)
  if (length(get_sign_values)>0){
    if (length(get_sign_values)==1){
      Gene_ID <- proteinGroups_norm$GeneID[i]
      Gene_name <- proteinGroups_norm$GeneName[i]
      Tissue <- names(proteinGroups_norm)[2+get_sign_values[1]]
      Category <- "Tissue enriched"
      new_row <- c(Gene_ID, Tissue, Gene_name, Category)
      Elevated_genes <- rbind(Elevated_genes, new_row)
    }
    else(
      if(length(get_sign_values)<8){
        for (j in seq(get_sign_values)){
          Gene_ID <- proteinGroups_norm$GeneID[i]
          Gene_name <- proteinGroups_norm$GeneName[i]
          Tissue <- names(proteinGroups_norm)[2+get_sign_values[j]]
          Category <- "Group enriched"
          new_row <- c(Gene_ID, Tissue, Gene_name, Category)
          Elevated_genes <- rbind(Elevated_genes, new_row)
        }
      }
    )
    
    
  }
  
}
Elevated_genes <- Elevated_genes[-1,]
```

GO over-representation test 
```{r}
get_GoTerm <- function(x) {
tissue<-x
genes <- Elevated_genes[Elevated_genes$Tissue==tissue,]$Gene.ID
total_genes <- proteinGroups_norm$GeneID[proteinGroups_norm[tissue]>=0]
totalgene.df <- bitr(total_genes, fromType = "ENSEMBL",
        toType = c("ENTREZID", "SYMBOL"),
        OrgDb = org.Hs.eg.db)
entrezid_total <- totalgene.df$ENTREZID
gene.df <- bitr(genes, fromType = "ENSEMBL",
        toType = c("ENTREZID", "SYMBOL"),
        OrgDb = org.Hs.eg.db)
entrezid <- gene.df$ENTREZID
ego <-NULL
ego <- enrichGO(gene          = entrezid,
                universe      = entrezid_total,
                OrgDb         = org.Hs.eg.db,
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
        readable      = TRUE)
ego_simplified <- simplify(ego)
return(ego_simplified@result)}


```

```{r}
results <- data.frame()
for (i in unique(Elevated_genes$Tissue)){
  tmp_get_go_term <- get_GoTerm(i)
  if (dim(tmp_get_go_term)[1]!=0){
  tmp <- cbind(tmp_get_go_term, i)
  results <- rbind(results, tmp)}
  
}
```

#Group analysis
```{r}
total_prot <- apply(apply(proteinGroups_table_tmp, 2, is.na), 2, function(x) dim(proteinGroups_table)[1] - sum(x))
total_prot_mat <- as.matrix(t(total_prot))

Group_enrich <- c()
Tissue_enrich <- c()
for (i in seq(names(total_prot))){
  Group_enrich[i] <- length(which(Elevated_genes$Tissue==names(total_prot)[i] & Elevated_genes$Category=="Group enriched"))
  Tissue_enrich[i] <- length(which(Elevated_genes$Tissue==names(total_prot)[i] & Elevated_genes$Category=="Tissue enriched"))
}
total_prot_mat<-rbind(total_prot_mat, Group_enrich)
total_prot_mat<-rbind(total_prot_mat, Tissue_enrich)


rownames(total_prot_mat) <- c("Group mixed", "Group enriched", "Organ enriched")
##Analysis of Group tissue specific
sort(total_prot_mat[2,]*100/total_prot_mat[1,])
sort(total_prot_mat[3,]*100/total_prot_mat[1,])

total_prot_mat[1,] <- total_prot_mat[1,] - total_prot_mat[2,] - total_prot_mat[3,]
colnames(total_prot_mat) <- c("Adipose Tissue", "Adrenal Gland", "Bone Marrow", "Brain", "Breast", "Colon", "Duodenum",
                              "Esophagus", "Fallopian Tube Oviduct", "Gall Bladder", "Heart", "Kidney", "Liver", "Lung", "Lymph Node",
                              "Ovary", "Pancreas", "Placenta", "Prostate", "Rectum", "Salivary Gland", "Small Intestine", "Smooth Muscle", "Spleen", "Stomach", "Testis", "Thyroid", "Tonsil", "Umbilical Artery", "Urinary Bladder", "Uterine Endometrium", "Vermiform Appendix")
par(mar=c(1,10,2,3))
barplot(total_prot_mat, horiz = TRUE, las = 1, cex.names =  1, cex.axis = 0.8,col = c("darkblue", "darkred", "yellow"), legend = TRUE, args.legend = list(cex=.7, xjust=0.5), xlab = "Number of mapped genes")




total_prot_mat["Group mixed","AdrenalGland"] <- 8924
plot <- as.data.frame(t(total_prot_mat))
plot$Tissues <- rownames(plot)
plot1 <- plot[,c(4,1,2,3)]
plot_long <- gather(plot1, Condition, Values, colnames(plot1[2]):colnames(plot1[4]), factor_key=TRUE)

plot_long$Condition <- factor(plot_long$Condition, levels = c("Organ enriched", "Group enriched","Group mixed")) 

plot_long$Tissues <- gsub(" ", "", plot_long$Tissues, perl=TRUE)
ggplot(plot_long, aes(x=Values, y=Tissues, 
                      fill=Condition)) + 
  geom_bar(stat="identity") +
  ylab("")+
  xlab("Identified canonical proteins")+
  theme_bw()+
  scale_x_continuous(breaks=seq(0, 12500, by = 2500))+
  scale_fill_manual(values=c("orange", "plum", "steelblue"))+
  #scale_fill_manual(values=c("orange", "plum", "steelblue"),
  #                  guide = guide_legend(
  #  direction = "horizontal",
  #  title.position = "left",
  #  label.position = "bottom",
  #  label.hjust = 0.5,
  #  label.vjust = 1,
  #  label.theme = element_text(angle = 90)
  #))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        #legend.position = c(0.91, 0.95),
        legend.key = element_blank(),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm'),
        legend.text=element_text(size=11))+
  ggtitle("")+coord_flip()

```