---
title: "GO analysis"
output: html_notebook
---


#Load libraries, DB and the main table

```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("clusterProfiler")

library(clusterProfiler)

BiocManager::install("org.Rn.eg.db")
library(org.Rn.eg.db)



proteinGroups_table <- read.table("/Users/ananth/Documents/MaxQuant_Bechmarking/Rat/Gene_distribution_in_organs-GeneNames-Median_bin_values-RAT.txt",
                                  header = TRUE,
                                  stringsAsFactors = FALSE)
#proteinGroups_table <- proteinGroups_table[,-c(35,36,37,38)]
```


The proteinGroups_table is a bin intensity matrix by organ

The following script  perform a median normalization across samples


```{r}
proteinGroups_table_tmp <- proteinGroups_table[,c(-1,-2)]
median_values <- apply(proteinGroups_table_tmp, 1, function(x) median(na.omit(x)))
proteinGroups_norm <- proteinGroups_table_tmp/median_values
proteinGroups_norm[is.na(proteinGroups_norm)] <- 0
proteinGroups_norm <- cbind(proteinGroups_table[,c(1,2)], proteinGroups_norm)
#rm(median_values, proteinGroups_table_tmp)
proteinGroups_mean <- apply(proteinGroups_norm[, -c(1,2)], 1, mean)

```

Each gene-tissue is classified in one of 3 categories: 
   - "Tissue enriched" at least 10-fold higher intesnity compared to average value and must be unique high value.
   - "Group enriched" at least 10-fold higher intesnity compared to average value in a group of 2-4 tissues compared to al other tissues
   - "Mixed" when not fulfil any of the above categories (they are discarded)

```{r}
Elevated_genes <- data.frame(
  "Gene ID" = NA,
  "Tissue" = NA,
  "Gene name" = NA,
  "Category" = NA
)

for (i in seq(dim(proteinGroups_norm)[1])){
  get_sign_values <- which(proteinGroups_norm[ i, c(-1,-2)] > proteinGroups_mean[i]*2)
  if (length(get_sign_values)>0){
    if (length(get_sign_values)==1){
      Gene_ID <- proteinGroups_norm$GeneID[i]
      Gene_name <- proteinGroups_norm$GeneName[i]
      Tissue <- names(proteinGroups_norm)[2+get_sign_values[1]]
      Category <- "Tissue enriched"
      new_row <- c(Gene_ID, Tissue, Gene_name, Category)
      Elevated_genes <- rbind(Elevated_genes, new_row)
    }
    else(
      if(length(get_sign_values)<5){
        for (j in seq(get_sign_values)){
          Gene_ID <- proteinGroups_norm$GeneID[i]
          Gene_name <- proteinGroups_norm$GeneName[i]
          Tissue <- names(proteinGroups_norm)[2+get_sign_values[j]]
          Category <- "Group enriched"
          new_row <- c(Gene_ID, Tissue, Gene_name, Category)
          Elevated_genes <- rbind(Elevated_genes, new_row)
        }
      }
    )
    
    
  }
  
}
Elevated_genes <- Elevated_genes[-1,]
```

GO over-representation test 
```{r}
get_GoTerm <- function(x) {
tissue<-x
genes <- Elevated_genes[Elevated_genes$Tissue==tissue,]$Gene.ID
total_genes <- proteinGroups_norm$GeneID[proteinGroups_norm[tissue]>=0]
totalgene.df <- bitr(total_genes, fromType = "ENSEMBL",
        toType = c("ENTREZID", "SYMBOL"),
        OrgDb = org.Rn.eg.db)
entrezid_total <- totalgene.df$ENTREZID
gene.df <- bitr(genes, fromType = "ENSEMBL",
        toType = c("ENTREZID", "SYMBOL"),
        OrgDb = org.Rn.eg.db)
entrezid <- gene.df$ENTREZID
ego <-NULL
ego <- enrichGO(gene          = entrezid,
                universe      = entrezid_total,
                OrgDb         = org.Rn.eg.db,
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
        readable      = TRUE)
ego_simplified <- simplify(ego)
return(ego_simplified@result)}


```

```{r}
results <- data.frame()
for (i in unique(Elevated_genes$Tissue)){
  tmp_get_go_term <- get_GoTerm(i)
  if (dim(tmp_get_go_term)[1]!=0){
  tmp <- cbind(tmp_get_go_term, i)
  results <- rbind(results, tmp)}
  
}
```

#Group analysis
```{r}
total_prot <- apply(apply(proteinGroups_table_tmp, 2, is.na), 2, function(x) dim(proteinGroups_table)[1] - sum(x))
total_prot_mat <- as.matrix(t(total_prot))

Group_enrich <- c()
Tissue_enrich <- c()
for (i in seq(names(total_prot))){
  Group_enrich[i] <- length(which(Elevated_genes$Tissue==names(total_prot)[i] & Elevated_genes$Category=="Group enriched"))
  Tissue_enrich[i] <- length(which(Elevated_genes$Tissue==names(total_prot)[i] & Elevated_genes$Category=="Tissue enriched"))
}
total_prot_mat<-rbind(total_prot_mat, Group_enrich)
total_prot_mat<-rbind(total_prot_mat, Tissue_enrich)


rownames(total_prot_mat) <- c("Group mixed", "Group enriched", "Organ enriched")
##Analysis of Group tissue specific
sort(total_prot_mat[2,]*100/total_prot_mat[1,])
sort(total_prot_mat[3,]*100/total_prot_mat[1,])

total_prot_mat[1,] <- total_prot_mat[1,] - total_prot_mat[2,] - total_prot_mat[3,]
colnames(total_prot_mat) <- c("Brain", "Heart", "Kidney", "Liver", "Lung", "Spinal Cord", "Tendon", "Testis")
par(mar=c(4,7,4,4))
barplot(total_prot_mat, horiz = TRUE, las = 1, cex.names =  0.7, cex.axis = 0.8,col = c("darkblue", "darkred", "yellow"), legend = TRUE, args.legend = list(cex=.7, xjust=0.6), xlab = "Number of mapped genes")

```